---
title: "Mini Project 2: Explore & Interrogate"
toc: true
toc-location: right
execute:
  echo: false
  warning: false
  message: false
params:
  number: 2
  weight: '9%'
  data_file: 'my_clean_data.csv'
  purpose: "An agent can generate a chart or a summary statistic in seconds. What it won't do on its own is ask a good question, or doubt the first pattern it finds. In this assignment you'll direct Claude Code to clean, count, and join two real data sources, but the questions you ask (and whether you trust what comes back) are on you."
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)

source(here::here("_common.R"))

# Read in rubric data
rubric <- read_csv(here::here('rubrics', 'mini2.csv'))
maxPoints <- rubric %>%
    filter(rating == "Excellent") %>%
    summarise(max = sum(maxPoints)) %>%
    pull(max)
```

```{r child = here::here("fragments", "mini.qmd")}
```

Your mission is to explore who actually "wins" the Olympics, and to find out whether that answer holds up once you look more carefully.

You'll work with two real data sources:

1. Athlete-level results from the [2016 Rio Summer Olympics](https://github.com/rfordatascience/tidytuesday/tree/main/data/2024/2024-08-06) (`olympics_2016.csv`)
2. Country population figures from [Our World in Data](https://ourworldindata.org/population-growth) (`population_2016.csv`).

Neither file is analysis-ready. The Olympics data is one row _per athlete per event_, which is not the same thing as one row _per medal_. The two files don't necessarily use the same country names either. And some entries in the Olympics data (like athletes competing as "Individual Olympic Athletes") don't correspond to any country at all. Getting past all of that and being able to still ask a good question and find a correct answer is the point of this assignment.

# 1. Get organized

Download and **unzip** [this template](../templates/mini2.zip) for your project, then open Positron to the unzipped template folder.

Once Positron opens, click on the `report.qmd` file. That is the primary file you will edit to conduct your analysis.

# 2. Document the data

Inside the `data` folder, there is a `README.md` file with some missing information. Click on that file and edit it to fill in the missing information for **both** data files:

- `olympics_2016.csv`: athlete-level results from the 2016 Rio Summer Olympics, filtered from the [TidyTuesday Olympics dataset](https://github.com/rfordatascience/tidytuesday/tree/main/data/2024/2024-08-06), which itself was compiled from [sports-reference.com](https://www.sports-reference.com/).
- `population_2016.csv`: 2016 country population figures from [Our World in Data](https://ourworldindata.org/population-growth).

# 3. Preview the data

Preview both files (e.g. using `head()` and `glimpse()`). A few things worth noticing before you plan anything:

- The Olympics data has one row **per athlete per event** -- look at a team sport like "Basketball" or a doubles event and see how many rows a single medal produces.
- Some `team` values have a `-1`, `-2`, etc. suffix (e.g. `"Czech Republic-1"`). Figure out what that suffix means before you decide whether to keep or strip it.
- Some `team` values don't match typical country names at all (e.g. `"Individual Olympic Athletes"`) -- these athletes aren't competing for a specific country.
- Compare how country names are spelled in each file. They don't all match.

# 4. Load the data

Have Claude Code read in both `olympics_2016.csv` and `population_2016.csv`. Pause here and think about your strategy before moving on -- don't let the agent start cleaning or joining yet.

# 5. Develop Your Strategy

Before you prompt the agent to write any code, **you** need a plan. None of the following are things an agent will reliably get right unless you tell it what to check:

**Write out your strategy** as a numbered list. It should address:

- **Counting medals correctly.** A country's medal count is not the number of rows it has in the data. What does one row actually represent, and how will you collapse rows into one count per medal?
- **Handling the team-number suffixes and non-country teams.** Which rows need a name fix, and which rows (if any) can't be joined to a country at all?
- **Joining to population.** The two files don't use identical country names. How will you find the mismatches, and how will you decide what to do with each one?
- **What you'll check afterward.** You won't know your join or your counts are right just because the code ran -- decide now what evidence would convince you.

**Deliverable:** A written strategy with numbered steps, in your `report.qmd` file, written *before* any cleaning or joining code. Be specific -- this is what you'll hand to the agent, and what you'll check its output against later.

# 6. Direct the Agent

Now prompt Claude Code to implement your strategy. Review what it produces as it goes -- read the code, read the diffs, and check intermediate output rather than accepting the final result all at once.

If the agent's first attempt gets something wrong -- counts rows instead of medals, joins on country name without checking for mismatches, silently drops "Individual Olympic Athletes" without telling you -- that's expected. Catching it is the point of the next step.

**Deliverable:** Clean, well-organized code in your report that produces (a) a correct medal count per country and (b) a joined data frame with population added.

::: {.callout-note}

**Note**: Claude Code should directly edit your `report.qmd` file.

:::

# 7. Interrogate

Getting a data frame that *looks* right is not the same as verifying it *is* right. This is the step that's graded most heavily in this assignment. Do not just ask Claude Code "is this correct?" -- it will generally say yes whether or not it actually checked anything. You need to design and interpret the checks yourself, even if the agent helps write the code that runs them.

Run and document at least these checks:

- **Medal count check**: look up the real, official [2016 Rio medal table on Wikipedia](https://en.wikipedia.org/wiki/2016_Summer_Olympics_medal_table) for at least 3-5 countries and compare it to what your code computed. If they don't match, figure out why before moving on.
- **Join check**: did every country with medals actually find a population match? List any that didn't, and say what you decided to do about each one and why.
- **The headline comparison**: for every country, compute its rank by total medal count, and separately its rank by medals *per capita* (medals divided by population). Show both rankings side by side (a table, a chart, or both) and write about which countries move the most between them -- and why that makes sense once you think about what "per capita" is actually measuring.

**Deliverable:** An Interrogate section in your report showing the checks you ran, their output, and what you concluded -- including anywhere you found and corrected a mistake in the agent's first attempt.

# 8. Explore Your Own Questions

Once your joined data frame is verified, ask **at least 2 more** research questions of your own, using either the Olympics data, the population-joined data, or both.

**Your 2 questions must use genuinely different chart types.** For example, "which country won the most medals" and "which sport awarded the most medals" are *both* ranking questions -- that doesn't count as 2 different types. Pick from (at least) these categories, and use no more than one question per category:

- **Ranking** -- which categories are highest/lowest (ordered bar chart)
- **Comparison** -- how specific groups compare to each other on a measure (grouped/dodged bars)
- **Distribution** -- how a numeric variable is spread across observations (histogram, boxplot, density)
- **Relationship** -- how two numeric variables relate to each other (scatterplot)
- **Composition** -- how a total breaks down into parts (stacked bar, waffle chart, pie chart)

For each question: state it, make the chart, and write at least one paragraph interpreting what you found.

**Deliverable:** At least 2 research questions, each with a chart of a different type from the list above and a written interpretation.

::: {.callout-note}

**Note**: this dataset is a single year, so a trend-over-time chart isn't available here -- see the bonus if you want to explore that.)

:::

# 9. Document Your Process

**Save your joined, cleaned data** as `my_clean_data.csv` in the project root folder.

**Write a brief reflection** addressing:

- What were the 2-3 biggest challenges in cleaning and joining this data?
- Where, if anywhere, did Claude Code's output need to be corrected, and how did you catch it?
- What specifically did you check to be confident your medal counts and join are correct?
- What would you do differently if you encountered similar data again?

**Deliverable:** Saved CSV file and written reflection (4-6 sentences).

# 10. Render and submit

```{r child = here::here("fragments", "mini-submit.qmd")}
```

# Bonus: Chase the trend (+5%)

The 2016-only data can't answer a trend-over-time question. For bonus credit, go get the [full historical Olympics dataset](https://github.com/rfordatascience/tidytuesday/tree/main/data/2024/2024-08-06) (1896-2016) and explore how a country's per-capita medal count has changed over time.

Fair warning: the historical data includes teams that no longer exist (the Soviet Union, East and West Germany, and others), and population data for defunct countries is much harder to track down cleanly than for 2016. You'll need to make and justify a real judgment call about how to handle this -- there isn't a clean, universally "correct" answer.

# Grading Rubric

### `r maxPoints` Total Points

```{r, echo=FALSE}
rubric %>%
  mutate(description = paste0("<b>", points, '</b><br>', description)) %>%
  select(-points) %>%
  spread(key = rating, value = description) %>%
  select(-category) %>%
  rename(Category = label) %>%
  arrange(order) %>%
  select(-order) %>%
  select(-maxPoints) %>%
  kable(format = 'html', escape = FALSE) %>%
  kable_styling(bootstrap_options = "striped") %>%
  column_spec(1, width = "16%") %>%
  column_spec(2:4, width = "28%")
```
