---
title: "Agentic Workflows & Version Control"
subtitle: "EMSE 4572/6572: Exploratory Data Analysis"
author: "John Paul Helveston"
institute: "The George Washington University | EMSE"
date: "September 2, 2026"
format:
  revealjs:
    theme: lexis.scss
    slide-number: c/t
    incremental: false
    code-line-numbers: false
    highlight-style: github
    width: 1600
    height: 900
    fig-width: 7
    fig-height: 4
execute:
  echo: true
  warning: false
  message: false
  freeze: false
---

```{r setup}
#| include: false
library(tidyverse)
library(countdown)
```

# {.inverse background-color="#121212"}

::: {style="text-align:center; margin-top:1em;"}
[Today]{.fancy .font200 .white}
:::

::: {.font120}
Last week you got the whole stack **installed**.<br>
Today you **use** it — for real, end to end.
:::

---

## The one loop for the whole semester {background-color="#121212" .inverse}

::: {.font140 style="text-align:center; margin-top:0.6em; line-height:1.6;"}
**prompt** [→]{.gray} **read the diff** [→]{.gray} **verify** [→]{.gray} **commit / push**
:::

. . .

<br>

::: {.font110}
The agent is **fast**, **confident**, and **sometimes wrong**.
:::

. . .

::: {.font110 .darkgreen}
Your job moves from *typing code* to **directing and checking it.**
:::

---

## Why version control + agents belong together

::: {.columns}
::: {.column width="50%"}
### Version control

- A **save point** you can return to
- A **cloud backup** of your work
- A record of *what changed and when*
:::

::: {.column width="50%"}
### Agentic workflow

- You **direct**, the agent **edits your files**
- Fast — and occasionally **wrong**
- You need a **safety net**
:::
:::

. . .

<br>

::: {.font120 style="text-align:center;"}
You **commit often** *precisely so* it's safe to let an agent<br>loose on your files. [The two are one workflow.]{.darkgreen}
:::

---

## What you'll actually do today

1. Make a repo and **push** to GitHub — with a button, no command line
2. Point **Claude Code** at some data and read what it does
3. Run the **whole loop**: ask → verify → commit → push
4. Watch the agent be **confidently wrong** — and *catch it*

. . .

::: {.font110 .darkgreen style="margin-top:0.8em;"}
Mostly **hands-on**. I talk a little; you practice a lot.
:::

# Part 1 — GitHub Desktop {background-color="#121212" .inverse}

## Three words, that's the whole model

::: {.columns}
::: {.column width="33%"}
### repo
A project folder that **tracks its own history**
:::

::: {.column width="33%"}
### commit
A **labeled save point** — "this is a state worth remembering"
:::

::: {.column width="33%"}
### push
Send your commits **up to GitHub** (the cloud copy)
:::
:::

. . .

<br>

::: {.font110}
No `git` in a terminal. Everything today is a **button in GitHub Desktop.**
:::

---

## {.inverse background-color="#121212"}

::: {.font200 .fancy style="text-align:center;"}
Practice 1
:::

::: {.font120}
**Make a repo and push to it.**

1. GitHub Desktop → **File › New Repository** → name it `eda-w2-practice`
2. **Publish** it to GitHub (make it private)
3. Add a line to `README.md`, then **Commit** (write a message!)
4. **Push** — then open it on github.com and confirm it's there
5. Change the README again → **Commit** → **Push** → look at **History**
:::

```{r}
#| echo: false
countdown(minutes = 15, warn_when = 60, top = 0, right = 0, font_size = "2em")
```

---

## What just happened

::: {.font110}
- Your work now exists in **two places** (your laptop + GitHub)
- Every commit is a point you can **come back to**
- That History tab is your **undo button** for the whole project
:::

. . .

<br>

::: {.font120 .darkgreen style="text-align:center;"}
Now let's give ourselves something worth undoing.
:::

# Part 2 — Meet Claude Code {background-color="#121212" .inverse}

## Claude Code, in Positron

::: {.font110}
- Open your **project folder** in Positron (not a stray file — the folder)
- Start **Claude Code** in the terminal pane
- You **describe** what you want; it **reads your files** and **proposes edits**
- You **read the diff** and decide: accept, or push back
:::

. . .

<br>

::: {.font110 .darkgreen}
Watch me do one first — then you'll do your own.
:::

::: {.footnote}
Prefer Python? You may use it. My demos are in R — but *you* still own the verification either way.
:::

---

## {.inverse background-color="#121212"}

::: {.font200 .fancy style="text-align:center;"}
Practice 2
:::

::: {.font120}
**Let Claude import some data — and read what it did.**

1. Drop `flights.csv` into your repo's `data/` folder
2. Ask Claude: *"Import `data/flights.csv` and give me a quick summary of what's in it."*
3. **Read the code it wrote.** Not just the summary — the actual code.
4. Tell a neighbor: *what did it do, and how do you know it's right?*
:::

```{r}
#| echo: false
countdown(minutes = 12, warn_when = 60, top = 0, right = 0, font_size = "2em")
```

---

## The habit to build now

::: {.font120}
The summary Claude prints is a **claim.**
:::

. . .

::: {.font120}
The code it wrote is the **evidence.**
:::

. . .

<br>

::: {.font130 .darkgreen style="text-align:center;"}
Always read the evidence, not just the claim.
:::

# Part 3 — The whole loop {background-color="#121212" .inverse}

## Ask → verify → commit → push

::: {.font110}
Now we string it together into one motion:
:::

::: {.font110}
1. **Ask** Claude to make a chart from the data
2. **Verify** it — do the numbers behind the chart match reality?
3. **Commit** the chart + code with a clear message
4. **Push** it to GitHub
:::

. . .

<br>

::: {.font110 .darkgreen}
This is the exact loop you'll run all semester. Everything else is details.
:::

---

## {.inverse background-color="#121212"}

::: {.font200 .fancy style="text-align:center;"}
Practice 3
:::

::: {.font120}
**Run the full loop yourself.**

1. Ask Claude to make a chart answering a question about `flights.csv`
2. **Verify**: pick one bar/point and check it against the data by hand
3. Does the chart *actually say what you asked?* If not, tell Claude what's wrong
4. **Commit** (chart + code) → **Push** to GitHub
:::

```{r}
#| echo: false
countdown(minutes = 20, warn_when = 120, top = 0, right = 0, font_size = "2em")
```

---

## You just did the whole thing

::: {.font120}
Directed an agent · read its work · **verified** it · versioned it · backed it up.
:::

. . .

<br>

::: {.font110}
That's the job. But we skipped the most important word.
:::

::: {.font160 .red style="text-align:center; margin-top:0.4em;"}
Verify.
:::

# Part 4 — Where it goes wrong {background-color="#121212" .inverse}

## The agent will be wrong *with confidence*

::: {.font110}
It won't say *"I'm not sure."* It will hand you a clean, plausible,<br>
**wrong** answer and move on. Common failure modes:
:::

::: {.columns}
::: {.column width="50%"}
- **Hallucinated** column names
- **Silently dropped** rows (`NA`s, bad joins)
- Wrong **join direction**
:::

::: {.column width="50%"}
- A chart that **misleads** (bad axis, wrong aggregation)
- A **made-up** number that "looks right"
- Code that isn't **reproducible**
:::
:::

. . .

::: {.font120 .red style="text-align:center; margin-top:0.5em;"}
Let me show you one live.
:::

---

## Your verification checklist {background-color="#EAE9EA"}

::: {.font110}
Run these every time — they're the skill this whole course is really about:

- Did **row / column counts** change the way I expected?
- Did I **read the code**, not just the prose summary?
- Did I **spot-check** a few values against the source file?
- Were rows **silently dropped** (NA handling, inner vs. left join)?
- Do **totals / aggregates** sanity-check?
- Is it **reproducible** — or did it hard-code an answer?
:::

::: {.footnote}
We'll come back to this list all semester. Screenshot it.
:::

---

## {.inverse background-color="#121212"}

::: {.font200 .fancy style="text-align:center;"}
Practice 4
:::

::: {.font120}
**Catch the bug.**

I've given Claude a task where its *first answer is wrong.* Your job is to
**find it** using the checklist.

1. Ask Claude to join `flights.csv` with `airlines.csv` and count flights per airline
2. **Something is off.** Verify before you believe it.
3. When you catch it, tell Claude *specifically* what's wrong and have it fix it
4. **Commit** the fix with a message that says what was broken
:::

```{r}
#| echo: false
countdown(minutes = 18, warn_when = 120, top = 0, right = 0, font_size = "2em")
```

---

## The catch is the whole job

::: {.font120}
An agent can write the code. [It cannot know if the answer is *true*.]{.red}
:::

. . .

::: {.font120 .darkgreen}
**That part is yours.** It's what you're actually being graded on.
:::

# Wrap-up {background-color="#121212" .inverse}

## Today, in one slide

::: {.font120 style="text-align:center; margin:0.4em 0; line-height:1.6;"}
**prompt** [→]{.gray} **read the diff** [→]{.gray} **verify** [→]{.gray} **commit / push**
:::

::: {.font110}
- You can **make a repo and push** with a button
- You can **direct Claude** to import, analyze, and chart
- You know the agent **lies with confidence** — and you have a checklist
- **R or Python** — your choice; the verification is the same job
:::

---

## Before next week

::: {.font110}
- Finish pushing today's work to your `eda-w2-practice` repo
- **Reading reflection:** read forward on tidy data + joins, try the light practice, reflect
- Next week: **Cleaning, Reshaping & Joining** — same loop, messier data
:::

. . .

<br>

::: {.font130 .fancy .darkgreen style="text-align:center;"}
Direct boldly. Verify relentlessly.
:::
